p-NLI: a Perturbation-Driven Visual Interrogation Tool for Analyzing and Interpreting Natural Language Inference Models

### 2018-03-21 paper/intro.tex
update reference

% (What is the high level motivation)
% What is the obsatcle for effectively apply machine learning model
With the recent advances in neural network based model, machine learning has gained unprecedented popularity, and its adaption permeated many fields of studies.
%
However, from researchers to practitioners, one often need to overcome many obstacles during the training, debugging, and tuning processes to unleash the full potential of these models.
%
Interpreting the internal mechanism and analyzing how predictions are made are critical to both the design and deployment of a given model.
More importantly, the ability to pinpoint where and how an error is made and come up with hypothesis for the cause of failure is the key to identify the limitation and improve upon existing models.
However, providing meaningful answers to these questions are extremely challenging and often regarded as impossible by many.

Recently, substantial research has been developed to combat the model interpretability challenge. Most notably, the works~\cite{} on interpretation of convolution neural network (CNN) model from both machine learning and visualization community has reveal the enormous potential and instant benefits of model interpretation, yet they also remind us how little we actually know regarding the inner mechanism of these deep neural network models.
%
Compare to image classification task, natural language processing systems have a number of innate restriction (e.g., the discrete nature of words) that makes interpreting them even more challenging. For example, \emph{feature visualization}~\cite{}, an often deployed technique for illustrating what kind of image feature a given part of the network (e.g., neuron, layer, or channel) captures, can not be easily generalized and applied for natural language due to the requirement of a continuous solution space (i.e., the image pixel values). Even though we can encode a word as an vector~\cite{}, the word embedding space is still a discretely defined space, where the interpolation between two vectors (i.e., two words) do not corresponds to explicit meaning.
%
In this work, we introduce a visualization system specific tailor to natural language model, specifical natural language inference, in which we address many domain specific challenges.

Many of the existing techniques study the model as an invariant object, where the model's behavior are recorded and analyzed in an offline fashion.
%
However, the exploration natural of these type of analysis often lead to \emph{what if ...} types of questions. What if I perturbed the current input? Will the prediction be stable? What if we change one of the key internal states of the model, how would the change affect the prediction. What if the current prediction is wrong, how and where should we change the model in order to produce the correct result? These type of dynamic queries is a natural way for us to gain understanding and come up with hypothesis of the model mechanisms by \emph{interrogating} the relationship between the input, internal states, and output prediction. Yet, we often need to tediously setup specialized experiment to obtain the answers.
%
In this work, we aim to provide quick and intuitive answers to these \emph{what if ...} questions. Instead of viewing the model as an invariant entitle, we employ a perturbation centric mentally to probe the model and examine how changes in one part of the pipeline affect others, which in term provide a new perspective to look at the model interpretation problem.

>>>>

% (What is the high level motivation)
% What is the obsatcle for effectively apply machine learning model
With the recent advances in neural network based model, machine learning has gained unprecedented popularity, and its adaption permeated many fields of studies.
%
However, from researchers to practitioners, one often need to overcome many obstacles during the training, debugging, and tuning processes to unleash the full potential of these models.
%
Interpreting the internal mechanism and analyzing how predictions are made are critical to both the design and deployment of a model.
More importantly, the ability to pinpoint where and how an error is made and come up with hypothesis for the cause of failure is the key to identify the limitation and improve upon existing models.
However, providing meaningful answers to these questions are extremely challenging and often regarded as impossible by many.

Recently, substantial research has been developed to combat the model interpretability challenge. Most notably, the works~\cite{SimonyanVedaldiZisserman2013, ZeilerFergus2014, YosinskiCluneNguyen2015, OlahMordvintsevSchubert2017, LiuShiLi2017, OlahSatyanarayanJohnson2018, BilalJourablooYe2018} on interpretation of convolution neural network (CNN) model from both machine learning and visualization community has reveal the enormous potential and instant benefits of model interpretation, yet they also remind us how little we actually know regarding the inner mechanism of these deep neural network models.
%
Compare to image classification task, natural language processing systems have a number of innate restriction (e.g., the discrete nature of words) that makes the road to interpretion more treacherous. For example, \emph{feature visualization}~\cite{OlahMordvintsevSchubert2017}, an often deployed technique for illustrating what kind of image feature a given part of the network (e.g., neuron, layer, or channel) captures, can not be easily generalized and applied for natural language due to the requirement of an interpretable continuous solution space, i.e., the image pixel values, in which human user can easily recongized (or the absent of) patterns or shapes. On the other hand, even though we have the ability to encode a word as an vector~\cite{MikolovSutskeverChen2013, PenningtonSocherManning2014}, the word embedding space is still a discretely defined space, where the interpolation between two vectors (i.e., two words) do not corresponds to explicit meaning.

These restrictions that prevent the application of the well-established interprability approach to natural language models motivated us to develop the proposed perturbation-driven technique.
%
In addition, many of the existing techniques study the model as an invariant object, where the model's behaviors are recorded and analyzed in an offline fashion.
%
However, the exploratory natural of the model interpretation analysis often lead to \emph{what if ...} types of questions. What if I perturbed the current input? Will the prediction be stable? What if we change one of the key internal states of the model, how would the modification affect the prediction. What if the current prediction is wrong, how and where should we change the model parameters in order to produce the correct result? These type of dynamic queries form a natural way for us to gain understanding and come up with hypothesis of the model mechanisms by \emph{interrogating} what a componenet do by perturb it and then observe the effects on other part of the system. Therotically, we  Yet, we often need to tediously setup specialized experiment to obtain the answers.

relationship between the input, internal states, and output prediction.
In this work, we aim to provide quick and intuitive answers to these \emph{what if ...} questions. Instead of viewing the model as an invariant entitle, we employ a perturbation centric mentally to probe the model and examine how changes in one part of the pipeline affect others, which in term provide a new perspective to look at the model interpretation problem.
